Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-24 paper.tex
almost conclude

To summarize, the primary contributions of this paper are:
\begin{itemize}
\item hypercoercions, a new space-efficient cast representation, which
  has a structurally recursive composition and a more compact memory
  representation.
\item a framework in Agda for proving the correctness of \lazyD{} cast
  represantations.
\item a formal proof that \lazyD\ hypercoercions respect the semantics
  of the \lazyD{} cast calculus.
\end{itemize}

@@

The $seq$ operator is defined in terms of coercion composition, which
is the key to compressing coercions and maintaining space efficiency.
A composition operator for coercions typically requires the target
type of the first coercion to match the source type of the second.
However, it is useful to relax this restriction for the D blame
tracking strategy.
%
We shall need optional blame labels, written $\ell$ , that range over
$\epsilon$ or $l$. The label is mandatory when the target
type of the first coercion does not match the source type of the second. 
Then we write composition of hypercoercions as $c_1
\fatsemi^\ell c_2$.  When both $c_1$ and $c_2$ are $\hyperCoercionI$,
their composition is also $\hyperCoercionI$.  If the head of $c_2$ is
a projection, the result is $c_2$.  Otherwise, we need a label to
build the projection. Since the head of $c_2$ is the no-op, its
source type must be a pretype. Thus we know $\ell$ must be a label and
we put it in the projection. When $c_1$ ends with a failure, the
composition is $c_1$ itself. In all the remaining cases, $c_1$ does
not end with a failure.  When $m_1$ and $m_2$ have different top
constructors ($m_1 \not\smile m_2$), the result is a failure coercion,
which needs a label. When $c_2$ starts with a projection, we blame the 
projection for casting a value to a shallowly inconsistent type. When $c_2$
starts with the no-op, we know $\ell$ must be a label, which is blamed for
composing shallowly inconsistent hypercoercions. The last two cases
compose $m_1$ and $m_2$ with the auxilliary function $m \fatsemi^\ell
m$, which assumes its inputs have the same top constructor. The
definition of $m \fatsemi^\ell m$ is a straightforward structural
recursion. Going back to the last two cases of $c_1 \fatsemi^\ell
c_2$. When $c_2$ starts with a projection, we do not know whether the
target type of $m_1$ matches the source of $m_2$. So we use the $l$ in the 
projection to  compose the middles $m_1 \fatsemi^l m_2$. When $c_2$ starts with 
the no-op, we compose the middles using $\ell$.

@@

This section presents our second contribution, a framework for proving
the correctness of cast representations, especially space efficient
ones. Section~\ref{sec:hypercoercion-correctness} applies this
framework to prove the correctness of the \lazyD{} hypercoercions.

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$
(Section~\ref{sec:framework:cek}). This machine is space efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly 
(Section~\ref{sec:framework:adt}).  We
conjecture that all cast representations defined in the literature
are instances of this ADT.

@@

%% After showing their cast is an instance of Lazy D Cast, users of our
%% frameworks can plug-in this cast representation to our space-efficient
%% abstract machine in subsection~\ref{ssec:framework:cek}, and prove
%% that this machine is equivalent to
%% \ineffCEKD\ (subsection~\ref{secc:framework:all-correct}), which
%% depends on lemmas in subsection~\ref{secc:framework:monoid-correct}.

@@

In this

>>>>

To summarize, the primary contributions of this paper are:
\begin{itemize}
\item hypercoercions, a new cast representation, which
  has a structurally recursive composition and a more compact memory
  representation.
\item a framework in Agda for proving the correctness of \lazyD{} cast
  represantations.
\item a formal proof that \lazyD\ hypercoercions respect the semantics
  of the \lazyD{} cast calculus.
\end{itemize}

@@

The $seq$ operator is defined in terms of coercion composition, which
is the key to compressing coercions and maintaining space-efficiency.
A composition operator for coercions typically requires the target
type of the first coercion to match the source type of the second.
However, it is useful to relax this restriction for the D blame
tracking strategy.
%
We shall need optional blame labels, written $\ell$ , that range over
$\epsilon$ or $l$. The label is mandatory when the target
type of the first coercion does not match the source type of the second. 
Then we write composition of hypercoercions as $c_1
\fatsemi^\ell c_2$.  When both $c_1$ and $c_2$ are $\hyperCoercionI$,
their composition is also $\hyperCoercionI$.  If the head of $c_2$ is
a projection, the result is $c_2$.  Otherwise, we need a label to
build the projection. Since the head of $c_2$ is the no-op, its
source type must be a pretype. Thus we know $\ell$ must be a label and
we put it in the projection. When $c_1$ ends with a failure, the
composition is $c_1$ itself. In all the remaining cases, $c_1$ does
not end with a failure.  When $m_1$ and $m_2$ have different top
constructors ($m_1 \not\smile m_2$), the result is a failure coercion,
which needs a label. When $c_2$ starts with a projection, we blame the 
projection for casting a value to a shallowly inconsistent type. When $c_2$
starts with the no-op, we know $\ell$ must be a label, which is blamed for
composing shallowly inconsistent hypercoercions. The last two cases
compose $m_1$ and $m_2$ with the auxilliary function $m \fatsemi^\ell
m$, which assumes its inputs have the same top constructor. The
definition of $m \fatsemi^\ell m$ is a straightforward structural
recursion. Going back to the last two cases of $c_1 \fatsemi^\ell
c_2$. When $c_2$ starts with a projection, we do not know whether the
target type of $m_1$ matches the source of $m_2$. So we use the $l$ in the 
projection to  compose the middles $m_1 \fatsemi^l m_2$. When $c_2$ starts with 
the no-op, we compose the middles using $\ell$.

@@

This section presents our second contribution, a framework for proving
the correctness of cast representations, especially space-efficient
ones. Section~\ref{sec:hypercoercion-correctness} applies this
framework to prove the correctness of the \lazyD{} hypercoercions.

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$
(Section~\ref{sec:framework:cek}). This machine is space-efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly 
(Section~\ref{sec:framework:adt}).  We
conjecture that all cast representations defined in the literature
are instances of this ADT.

@@

@@

In this paper we presented a new cast representation, hypercoercions, in two 
flavors, Lazy D and Lazy UD. Hypercoercions have a structurally recursive 
composition, which is a win over coercions in normal form for mechanized 
metatheory. What is more, hypercoercions have a more compact memory 
representation. We prove that $S(H)$, an abstract machine that uses Lazy D 
hypercoercions, is equivalent to $D$, a semantics of Lazy D Cast Calculus.

We also present a framework for proving the correctness of Lazy D Cast ADT. 
If $C$ implements the Lazy D Cast ADT, $S(C)$ would be equivalent to $D$.

\todo[inline]{related works. -KC}
